Model to study co-evolution of trade an culture

### 2015-04-01 article.tex
minor modifications

\subsection{Modelling of Trade}

@@

Among the various works done with this objective, few are studying the interactions between economy and cultural change in Roman empire~\cite{terpstra_trade_2011,kessler_organization_2007,scheidel_model_2007}. 
An analysis based on historical documents has revealed the importance of economical and social institutions for the efficiency of trades of grains in the early Roman Empire~\cite{kessler_organization_2007}. The same type of analysis has also revealed the importance of a number of key factors () for the growth of income and population in the Roman Italy~\cite{scheidel_model_2007}. Finally, a micro-economic model has determined the importance of reputation and threat of expulsion from a trade network as mechanisms to maintain the overall efficiency of the trade network~\cite{terpstra_trade_2011}. These works therefore reveal the strong relation between trade and cultural change in the Roman empire.

However, a limitation of the previous works is the impossibility to test the conclusions reached by the authors. The models proposed reflects only the view of the authors on the data they have analyse. A new type of model would be necessary to go beyond the interpretation, and allow the verification of the conclusions reached. These models should be able to generate data which is then qualitatively compared to the real historical data. By this mean, if a factor is really important in the economical network, its removal from the model will lead to data qualitatively different from the historical data. The Multi-Agent Simulations (MAS) offer a way to produce this kind of model~\cite{lake_trends_2014}.

>>>>

\subsection{Models of Trade}

@@

Among the various works done with this objective, few are studying the interactions between economy and cultural change in the Roman empire~\cite{terpstra_trade_2011,kessler_organization_2007,scheidel_model_2007}. 
An analysis based on historical documents has revealed the importance of economical and social institutions for the efficiency of trades of grains in the early Roman Empire~\cite{kessler_organization_2007}. The same type of analysis has also revealed the importance of a number of key factors (such as commercial development, population growth and geographical mobility) for the growth of income and population in the Roman Italy~\cite{scheidel_model_2007}. Finally, a micro-economic model has determined the importance of reputation and threat of expulsion from a trade network as mechanisms to maintain the overall efficiency of the trade network~\cite{terpstra_trade_2011}. These works therefore reveal the strong relation between trade and cultural change in the Roman empire.

However, a limitation of the previous works is the impossibility to test the conclusions reached by the authors. The models proposed reflects only the view of the authors on the data analysed. A new type of model would be necessary to go beyond the interpretation, and allow the verification of the conclusions reached. These models should be able to generate data which is then qualitatively compared to the real historical data. By this mean, if a factor is really important in the economical network, its removal from the model will lead to data qualitatively different from the historical data. The Agent Based Models (ABM) methodology offers a way to produce this kind of model~\cite{lake_trends_2014}.
